---
title: "RVision Mini-Gallery"
author: "Bo Yuan"
date: "11/5/2019"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_depth: 4
    toc_float: 
      collapsed: true
      smooth_scroll: false
---

This document serves as a quick demonstration of some of the useful codes I've been using frequently for R plotting. All dataset used is built in R, and there is no need to import external data (web-linkage, literature, etc.) to run the script. 

The script should run smoothly without error. If there are any problems, please report to yuanbo.faith@gmail.com.

More plots on the way.  


```{r, message=F, warning=F}
library(ggplot2)
library(tidyr)
library(dplyr)
library(stringr)
library(rebus)
library(gridExtra)
library(cowplot)
library(ggrepel)
library(ggExtra)
library(RColorBrewer)
library(viridis)
library(nycflights13)
library(MASS)
library(leaflet) # map drawing
library(ggthemes) # for easeir theme setting
library(wordcloud)
```


# Scatterplot

## Old Faithful Geyser
```{r, message=F, warning=F}
head(faithful)

p = ggplot(faithful, 
           aes(x = waiting, y = eruptions)) + 
  geom_point(alpha = .5,
             size = 3,
             position = "jitter") +
  theme_classic() +
  theme(title = element_text(face = "bold"))

p   

ggMarginal(p) # add marginal density plot
```



## Iris flowers
### Plot 1

```{r, message=F, warning=F}
ggplot(iris, 
       aes(Sepal.Length, y = Sepal.Width, color = Species)) + 
  geom_count(alpha = .6) + # use point size to show data density (overcome overlaping)
  theme_classic() +
  labs(title = "Edgar Anderson's Iris Data") +
  theme(title = element_text(face = "bold"))
```


### Plot 2

```{r, message=F, warning=F}
head(iris)
ggplot(iris, 
       aes(x = Sepal.Length, y = Sepal.Width, 
           color = Petal.Length, size = Petal.Width)) + 
  geom_point(position = position_jitter(.2, .2), alpha = .6) + # add small random noise to dots' position to avoid overlap
  geom_smooth(method = "lm", color = "black", se = F, size = .5) +
  facet_wrap(~Species) + # facet based on species 
  
  # polish up
  theme_bw() +
  scale_color_continuous(low = "steelblue", high = "black") + # control colorbar for Petal.Length
  theme(title = element_text(face = "bold"),
        axis.title =  element_text(color = "firebrick", face = "bold"),
        axis.text = element_text(color = "firebrick"),
        strip.background = element_blank(), 
        strip.text = element_text(face = "bold", color = "firebrick", size = 12)) +
  labs(title = "Edgar Anderson's Iris Data")
```


## Diamonds dataset
```{r, message=F, warning=F}
head(diamonds)
dim(diamonds)

ggplot(diamonds, 
       aes(x = carat, y = price, color = cut)) + 
  geom_point(alpha = .4, shape = ".") + # "shape = "." and transparency set to .4 decrease overlap 
  geom_smooth(method = "lm", se = F, size = .6, linetype = "dashed") + # regression
  
  # polish up
  scale_color_brewer(palette = "Set1") + 
  theme_bw() + 
  theme(legend.position = c(.8, .2), 
        title = element_text(face = "bold")) +
  coord_cartesian(xlim = c(0, 4), ylim = c(0, 20000)) + # zoom in 
  labs(title = "Prices of 50,000 round cut diamonds")
```


## Cars dataset
```{r, fig.width=10, fig.height=8}
p = ggplot(mtcars, 
           aes(x = wt, y = mpg, 
               color = factor(cyl), shape = factor(am))) + # use of "factor()" change cyl and am from continuous to discrete variables; particularly important since "shape" is discrete
  geom_point() +
  geom_smooth(aes(group = cyl), method = "lm", se = F, linetype = "dashed") + # regression line. specification of "aes(group = cyl)" ensures grouping based on cylinder (cyl) number alone; if not specified, inheriting aesthetic from above, regression is performed  based on combinations of cyl and auto/manual transmission (am)
  geom_text_repel(aes(label = rownames(mtcars)), size = 3) + # add car names, avoiding overlap
  
  # polish up
  theme_classic() + 
  theme(legend.position = c(.8, .7), 
        title = element_text(face = "bold")) +
  stat_ellipse(aes(group = cyl), linetype = "dotted") + # add distribution ellipse
  labs(x = "Weight (1000 lbs)", y = "Miles/(US) gallon")
p  

ggMarginal(p, type = "density", size = 10, groupColour = T, groupFill = T, alpha = .2) # add marginal density plot
```



## Mammals dataset
```{r, message=F, warning=F}
ggplot(mammals, 
       aes(x = body, y = brain)) + geom_point() +
  geom_smooth(method = "lm", fill = "firebrick", color = "firebrick", alpha = .2, size = .5) +
  scale_x_log10() + 
  scale_y_log10() + # log transform axis
  
  # polish up
  annotation_logticks() +  # add log10 marks
  theme_bw() +
  labs(title = "Attributes of Animals data") +
  theme(title = element_text(face = "bold"))
```


## New york flight dataset
```{r, message=F, warning=F, fig.height=9, fig.width=10}
theme_set(theme_bw() + 
            theme(title = element_text(colour = "black", face = "bold"),
                  axis.text = element_text(colour = "black", face = "bold"),
                  legend.text = element_text(colour = "black", face = "bold"),
                  legend.title = element_text(colour = "black", face = "bold")))

planes %>% 
  ggplot(aes(x = year, y = seats, color = engine)) +
  geom_count(alpha = .4, 
             position = position_jitter(0, 10)) +
  scale_x_continuous(breaks = seq(1960, 2010, by = 10)) +
  scale_color_brewer(palette = "Dark2", "Engine type") +
  scale_size(range = c(4, 14), "Dot density") +
  theme(legend.position = c(.2, .7)) +
  labs(title = "Plane metadata")
```



# Lines plot

## Loblolly dataset
```{r, message=F, warning=F}
# The growth of Loblolly pine trees
p1 = Loblolly %>%
  ggplot(aes(x = age, y = height, color = Seed)) +
  geom_point() +
  geom_line() +
  theme(legend.position = "")


# Set up different colors
color.func = brewer.pal(11, "Spectral") %>% colorRampPalette() # divide the "Spectral" 1 colors into n colors;
seed.color = color.func(Loblolly$Seed %>% n_distinct()) %>% # n = number of seeds
  sample() # randomly shuffle the color orders so that similar colors may be displayed apart

p2 = p1 + scale_color_manual(values = seed.color) 

grid.arrange(p1, p2, nrow = 1)
```


## Indometh dataset
### Spaghetti plot 
```{r, message=F, warning=F}
# spaghetti plot 
p1 = 
  Indometh %>% 
  ggplot(aes(x = time, y = conc, color = Subject)) +
  geom_point(position = position_dodge(.1)) +  # add random noise to point coordinates to avoid overlap
  geom_line(position = position_dodge(.1)) +  # add random noise to line coordinates to avoid overlap, in line with points
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  theme(legend.position = "NA",
        panel.grid = element_blank()) 


p2 = p1 + scale_y_log10() +
  labs(y = "conc. logarithmically transformed") + # log transform y-axis
  annotation_logticks() # add scale ticks


ggdraw() +
  draw_plot(p1) +
  draw_plot(p2, x = .3, y = .3, width = .6, height = .6)
```

### Spaghetti plot with trend line
```{r, message=F, warning=F}
# Create dataset containing averaged trend line

# Create sumamry dataset of the trending line
Indometh.summary = 
  Indometh %>% 
  group_by(time) %>%
  summarise(conc.mean = mean(conc),
            conc.sd = sd(conc),
            conc.max = max(conc),
            conc.min = min(conc))
Indometh.summary


p1 = Indometh %>%
  ggplot(aes(x = time)) + # x = time shared by all following points and line aesthetic, so specified here
  geom_point(aes(y = conc, color = Subject)) + # "conc" and "Subject" not shared by Indometh.summary, so specified here instead of upper line
  geom_line (aes(y = conc, color = Subject)) +
  scale_color_brewer(palette = "Set2") +
  
  # Add averaged trending line
  geom_line(data = Indometh.summary,
            aes(y = conc.mean), color = "black") +
  
  geom_errorbar(data = Indometh.summary,
                aes(ymin = conc.mean - conc.sd, 
                    ymax = conc.mean + conc.sd)) +
  
  geom_point(data = Indometh.summary,
             aes(y = conc.mean), 
             shape = 21, fill = "white", size = 3) + # most top layer above the lines and errorbars
  theme_bw() +
  theme(legend.position = "NA") 


p2 = p1 + coord_cartesian(xlim = c(0, 2)) +
  labs(title = "Zoom-in view\n for denser time intervals")


ggdraw() +
  draw_plot(p1) +
  draw_plot(p2, x = .3, y = .3, width = .6, height = .6)
```


## Airquality dataset
```{r, message=F, warning=F}
head(airquality)

airquality1 = airquality %>% 
  arrange(Month, Day) %>% 
  mutate(Time = 1:nrow(airquality) )

high.temp = airquality1 %>% filter(Temp >= 90)  

airquality1 %>%
  ggplot(aes(x = Time, y = Temp,
             color = factor(Month))) +
  geom_line(aes(group = 1)) +
  geom_point(shape = 21, fill = "white", stroke = 1) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        panel.grid = element_blank(),
        legend.position = c(.6, .05),
        legend.direction = "horizontal") +
  
  # rectangle highlighting high-termperature region
  geom_rect(aes(xmin = Time-1, xmax =Time, ymin = 90, ymax = 98), 
            # many small rectangles combined together...one rectangle for each row
            color = NA, fill = "firebrick", alpha = .1) +
  
  # highlight high-temperature threshold line
  geom_rect(aes(xmin = Time-1, xmax = Time, ymin = 90, ymax = 90), 
            # many small segments combined together...one segment for each row
            color = "firebrick", size = .2) +
  
  # Text annotation
  annotate("text", x = 20, y = 97, label = "HIGH HEAT ZONE",
           fontface = "bold", color ="firebrick") +
  
  # highlight high-termperature points with red fill
  geom_point(data = high.temp, 
             shape = 21, fill = "firebrick", size = 3, show.legend = F) +
  # new bigger points cover prior points; x, y coordinates inherit upper lines
  
  # Text label for high temperature           
  geom_text_repel(data = high.temp,
                  aes(label = Temp), # x, y coordinates inherit upper lines
                  size = 3, color = "firebrick") 

```



# Bar plots

## Car dataset
```{r, message=F, warning=F}
# convert from matrix to data frame or tibble
mtcars.t = as_tibble(mtcars)
mtcars.t$cars = rownames(mtcars)


# Arrange cars in order of horsepower (hp)
mtcars.t = mtcars.t %>% arrange(hp)
cars.ordered = mtcars.t$cars # vector of cars in order of horsepower

p.bar = mtcars.t %>% 
  mutate(cars = factor(cars, levels = cars.ordered, ordered = T)) %>%
  ggplot(aes(x = cars, y = hp, fill = factor(cyl))) +
  geom_bar(stat = "identity") +
  coord_flip() +
  
  # polish up
  
  theme_minimal() +
  theme(legend.position = c(.8, .2),
        axis.text = element_text(color = "black")) +
  
  # Add pattern/trending line
  geom_point(show.legend = F) +
  geom_line(aes(group = NA)) # group = "" or group = 1...both work too!
p.bar
```


```{r, fig.width=7, fig.height=7}
# polar plot
p.polar = p.bar + coord_polar() +
  
  # polish up
  theme(legend.position = c(.5, .2),
        legend.direction = "horizontal",
        axis.text.y = element_blank(),
        axis.title = element_blank())  +
  
  annotate(geom = "text", x = .5, y = .5, # put x, y coordinates here otherwise return an error
           label = "Spinning\nhorsepower", 
           color = "white", fontface = "bold", size = 5) +
  scale_fill_brewer(palette = "Dark2") 
p.polar
```


```{r, fig.width=8.5, fig.height=8.5}
# Rotated car names
text.angle = 90 - 360 * (1:nrow(mtcars) - .5)/nrow(mtcars) # rotate cars name together with bars
text.angle = ifelse(text.angle < -90, 180 + text.angle, text.angle) # control flipping direction
p.polar +
  geom_text(aes(y  =hp + 80, label = cars), angle = text.angle) +
  theme(axis.text = element_blank(),
        legend.position = c(.5, 0.15))
```



# Histogram 


## Iris dataset
```{r, message=F, warning=F}
iris.tidy = gather(iris, 1:4, # gather the first four variables into one variable (named as "measure)
                   key = measure, # the new variable "measure" contains the prior four variable names
                   value = dimension) # "dimension" contains all numeric values associated with the prior four variables

p1 = ggplot(iris.tidy, 
            aes(x = dimension, fill = Species)) + 
  geom_histogram(binwidth = .2, position = "dodge") +
  facet_wrap(~measure, nrow = 1,
             scales = "free_x") + # independent x-axis for each facet
  
  scale_fill_brewer(palette = "Accent") + 
  theme_bw() + 
  theme(legend.position = "bottom") +
  labs(title = "Histogram")
```


```{r, message=F, warning=F}
p2 = ggplot(iris.tidy, 
            aes(x = dimension, 
                color = Species, # "color": for the color of the boarder line 
                fill = Species)) + # "fill": for the color of the filling area
  geom_density(alpha = .3) + # adjust the filling transparency
  facet_wrap(~measure, nrow = 1, 
             scales = "free_x") + # independent x-axis for each facet
  
  scale_color_brewer(palette = "Accent") +
  scale_fill_brewer(palette = "Accent") +
  theme_bw() + 
  theme(legend.position = "bottom") +
  labs(title = "Density plot")
```


```{r, fig.width=10, fig.height=10}
grid.arrange(p1, p2, nrow = 2) # combine
```


## Rock dataset
```{r, message=F, warning=F}
head(rock)

# Check histogram for all four variables
gather(rock, 1:4, # gather all four variables into one variable named "measure"
       key = measure, # the new variable "measure" contains all prior four variable names
       value = values # "values" contains all numeric values associated with prior four variables
       ) %>% # the pipeline feeds the gathered dataset into the 1st argument position of ggplot and streamlines programming
  ggplot(aes(x = values)) + 
  geom_histogram(fill = "Steelblue") +
  facet_wrap(~measure, # facet based on "measure", i.e., the prior four variables
             scales = "free") + # independent x-y scale for all four faceted plots
  theme_minimal() +
  theme(strip.text = element_text(color = "firebrick", face = "bold"),
        panel.grid = element_line(color = "firebrick", size = .2))
```


## Diamonds dataset
```{r, message=F, warning=F}
ggplot(diamonds, aes(x = cut, fill = clarity)) + 
  geom_histogram(stat = "count", position = "fill") +
  theme_minimal()

```



# Lollipop plot

## Cars dataset

```{r, message=F, warning=F}
# convert from matrix to data frame or tibble
mtcars.t = as_tibble(mtcars)
mtcars.t$cars = rownames(mtcars)
mtcars.t

# Arrange cars in order of horsepower (hp)
mtcars.t = mtcars.t %>% arrange(hp)
cars.ordered = mtcars.t$cars # vector of cars in order of horsepower

mtcars.t.ordered = mtcars.t %>% 
  mutate(cars = factor(cars, levels = cars.ordered, ordered = T))
```


```{r, message=F, warning=F}
p1 = mtcars.t.ordered %>%
ggplot(aes(x = cars, y = hp, color = factor(am))) + 
  geom_point(aes(size = factor(cyl))) + 
  # Do not specify point size in upper ggplot() line; 
  # otherwise geom_segment lines width will also be dependent on cylinder number, which is less desirable 
  
  geom_segment(aes(x = cars, xend = cars, y = 0, yend = hp)) +
  coord_flip() +
  
  # polish up
  scale_color_brewer(palette = "Set2") +
  theme_minimal() +
  theme(legend.position = c(.8, .3),
        axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black", face = "bold"))
p1
```

----

*Ignore warning "Using size for a discrete variable is not advised". In this case, treating cylinder number as factor a discrete variable is more desirable than otherwise as numeric continuous variable.*

**Key in ordered plotting is to convert the associated variables (cars in this case) into: ORDERED FACTOR.** 


```{r, fig.width=9, fig.height=9}
# polar plot
text.angle = 90 - (1:nrow(mtcars) - .5) / nrow(mtcars) * 360
text.angle = ifelse(text.angle < -90, 180 + text.angle, text.angle)

p1 + coord_polar() +
  geom_text(aes(y = hp + 100, label = cars),
            angle = text.angle,
            show.legend = F) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        legend.position = c(.5, .15),
        legend.direction = "horizontal",
        panel.grid = element_blank())

```


# Maps
## world map

```{r, fig.width=10, fig.height=8}
my.world = map_data("world")
dim(my.world); head(my.world); tail(my.world)

n.country = my.world$region %>% n_distinct() 
color.func = brewer.pal(12, "Paired") %>% colorRampPalette() # Divide "Paired" colors into more colors
country.colors = color.func(n.country) # randomly shuffle color order; otherwise adjacent regions will be of similar colors

worldmap = ggplot() + 
  geom_polygon(data = my.world,
               aes(x = long, y = lat, # longitude vs. latitude
                   group = group, # dots for same group connected together
                   fill = region), # fill color by region (country)
               color = "white" # country boarder
               ) +
  
  # polish up
  guides(fill=FALSE) + # not show legend
  coord_fixed(1.2) + # keep aspect ratio the same
  scale_fill_manual(values = country.colors) +
  scale_x_continuous(breaks = seq(-180, 180, by = 60),name = "Longitude") +
  scale_y_continuous(breaks = seq(-90, 90, by = 30), name = "Latitude") +
  theme_minimal() 
  
worldmap
```


## BRICS VS NATO

Disclaim: this plot does not display any political views. It is only used for communicating plotting techniques. 

```{r, message=F, warning=F}
# Brics countries
Brics = c("Brazil", "Russia", "India", "China", "South Africa")
  # source: https://www.investopedia.com/terms/b/brics.asp
Brics.col = brewer.pal(length(Brics), "Reds")  # color for Bricks countries
names(Brics.col) = Brics; Brics.col # create named vectors


# NATO countries
NATO = c("Albania", "Belgium", "Bulgaria", "Canada", "Croatia", "Czech Republic", "Denmark", "Estonia", "France", "Germany", "Greece", "Hungary", "Iceland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Montenegro", "Netherlands", "Norway", "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Turkey", "UK", "USA")
# source : https://www.nti.org/learn/treaties-and-regimes/north-atlantic-treaty-organization-nato/

NATO %in% (map_data("world")$region %>% unique()) # check all NATO countries are listed in map_data 

# [3:9] to remove the less visible light blue colors
NATO.col.func = c(brewer.pal(9, "Blues")[3:9]) %>% colorRampPalette() 

NATO.col = NATO.col.func(n = length(NATO)) 
names(NATO.col) = NATO; NATO.col # create named vectors


# Combine Bricks with NATO
country = c(Brics, NATO)
country.colors = c(Brics.col, NATO.col)
```


```{r, fig.width=12, fig.height=8}
# Plot
p.world = ggplot(data = map_data("world"), # add global background
       aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "grey", color = "white") +
  coord_fixed(1.2) +
  theme_minimal() +
  
  geom_polygon(data = map_data("world", region = country), # add NATO & BRICS countries
               aes(fill = region), color = "white") +
  scale_fill_manual(values = country.colors) +
  
  # extra polish up
  labs(title = "NATO (blues) vs BRICS (reds)") +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        title = element_text(face = "bold")) +
  guides(fill = guide_legend(nrow = 4)) + # legend row number
  scale_x_continuous(breaks = seq(-180, 180, by = 60),name = "Longitude") +
  scale_y_continuous(breaks = seq(-90, 90, by = 30), name = "Latitude") 
  
p.world
```


```{r, message=F, warning=F}
# europe zoom in (as insert)
p.euro = p.world + 
  coord_fixed(1.2, xlim = c(-25, 50), ylim = c(30, 70)) + 
  # theme, labs and scale overwite those of p.world 
  theme(panel.border = element_rect(fill = NA, color = "black"), 
        panel.grid = element_blank(),
        title = element_text(face = "bold"),
        axis.title = element_blank(),
        legend.position = "") +
  labs(title = "European NATO zoomed-in") + 
  scale_x_continuous(breaks = seq(-25, 50, by = 25)) +
  scale_y_continuous(breaks = seq(30, 70, by = 10)) 

p.euro
```


```{r, fig.width=12, fig.height=10}
ggdraw() +
  draw_plot(p.world) +
  draw_plot(p.euro, x = 0.05, y = -.06, width = .27)
```


## US crime dataset
```{r, message=F, warning=F}
USArrests.2 = USArrests %>% 
  mutate(region = rownames(USArrests) %>% # add "region" column of state names 
           tolower()) # all states in small letters (to merge with "state" dataset)

 # these two states are not in the "state" mainland dataset
USArrests.2[!USArrests.2$region %in% map_data("state")$region, ]
  
USArrests.2 = USArrests.2 %>% 
  inner_join(map_data("state"), by = "region") # augment with US geographical data

USArrests.2 %>% 
  ggplot() +
  theme_map() +
  geom_polygon(aes(x = long, y = lat, group = group, fill = Murder), 
               color = "white", size = 1) + # set up state border line color and width
  
  # polish up
  coord_fixed(1.4) +
  labs(title = "Murder per 100K residents, USA, 1973") +
  scale_fill_viridis(option = "magma", direction = -1) +  # direction: reverse the color scale
  theme(legend.position = "bottom") +
  guides(fill = guide_colourbar(barwidth = 20, direction = "horizontal")) +
  theme(plot.title = element_text(hjust = .5, size = 14, face = "bold"))
```


## Database map
```{r, fig.width=8, fig.height=7}
library(leaflet)
# Background 1: NASA
leaflet() %>% 
  addTiles() %>% 
  setView( lng = 2.34, lat = 48.85, zoom = 5 ) %>% 
  addProviderTiles("NASAGIBS.ViirsEarthAtNight2012")


# Background 2: World Imagery
leaflet() %>% 
  addTiles() %>% 
  setView( lng = 2.34, lat = 48.85, zoom = 3 ) %>% 
  addProviderTiles("Esri.WorldImagery")
```


## US mainland airports
```{r, fig.width=10, fig.height=8}
library(stringr)
airport.US.mainland = airports %>%
  filter(tzone %>% str_detect("America"),
         lon < -50, lat < 50)

# select EWR airport
airport.EWR = airports %>% filter(faa == "EWR")
```

**Remove EWR from mainland airport dataset (otherwise geom_curve connection would report and error, as the same EWR point can't connect to itself)**

```{r, fig.width=10, fig.height=8}

airport.US.mainland = airport.US.mainland %>% anti_join(airport.EWR)


ggplot(data = map_data("usa") ) + # US map background
  geom_polygon(aes(x = long, y = lat, group = group),
               fill = "black") +
  coord_fixed(1.5) +
  theme_map() + # from ggthemes package
  
  # add airports
  geom_point(data = airport.US.mainland,
             aes(x = lon, y = lat),
             color = "tomato", shape = ".") +
  
  
  # connect EWR airport with all other airports
  geom_curve(data = airport.US.mainland,
             aes(x = airport.EWR$lon, xend = lon,
                 y = airport.EWR$lat, yend = lat),
             alpha = .1, color = "turquoise2",  
             curvature = .1) +
  annotate(geom = "Text", label = "EWR", color = "Black", fontface = "bold",
           x = airport.EWR$lon + 1.8, y = airport.EWR$lat-.8)

```


# Word cloud
```{r, message=F, warning=F}
library(nycflights13)
library(wordcloud)
plane.makers = planes %>% dplyr::select(manufacturer) %>% 
  table() # count occurrence frequency

set.seed(1234)
par(bg="black")  # set background
wordcloud(words = names(plane.makers), 
          freq = plane.makers %>% sqrt(), # reduce frequency difference
          min.freq = 1, random.order = F, rot.per = .1,
          colors = terrain.colors(8)) 

```

